---
title: "Relative Abundance Analysis from Kraken2 Reports"
author: "Enhanced Version"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: show
---

```{r setup, message=FALSE, warning=FALSE}
# =============================================================================
# Relative Abundance Analysis from Kraken2 Reports
# For Nanopore Sequencing Data
# Includes Species, Genus, and Family level analyses
# 
# UPDATED: 
#   - Excludes Homo sapiens reads and recalculates relative abundance
#   - Ensures all bars sum to 100%
#   - Includes sample renaming functionality
#   - Fixed sample ordering
#   - Enhanced color palette with more distinct colors
#   - Improved y-axis with more tick marks
# =============================================================================

# Load required libraries
library(tidyverse)
library(RColorBrewer)
library(scales)

# -----------------------------------------------------------------------------
# CONFIGURATION - Update this path to your data directory
# -----------------------------------------------------------------------------
data_dir <- "/Volumes/Alice/1_METHOD_DEVELOPMENT/Nanopore_runs"

# Number of top taxa to display (increase to reduce "Other" category)
n_top_taxa <- 15

# Maximum character width for labels before wrapping (adjust as needed)
label_wrap_width <- 25

# -----------------------------------------------------------------------------
# SAMPLE NAME MAPPING - Customize your sample names here
# Format: "barcodeXX" = "Your Custom Name"
# -----------------------------------------------------------------------------
sample_name_mapping <- c(
  "barcode01" = "Chicken 1-Non-enriched",
  "barcode02" = "Chicken 1-BPW+RVS",
  "barcode03" = "Chicken 1-Bolton base+CAT",
  "barcode04" = "Chicken 2-Non-enriched",
  "barcode05" = "Chicken 2-BPW+RVS",
  "barcode06" = "Chicken 2-Bolton base+CAT"
)

# Set to TRUE to use custom names, FALSE to use default barcode names
use_custom_names <- TRUE

# -----------------------------------------------------------------------------
# SAMPLE ORDER - Define the exact order samples appear on plots
# This MUST match your custom names exactly
# -----------------------------------------------------------------------------
sample_order_custom <- c(
  "Chicken 1-Non-enriched",
  "Chicken 1-BPW+RVS",
  "Chicken 1-Bolton base+CAT",
  "Chicken 2-Non-enriched",
  "Chicken 2-BPW+RVS",
  "Chicken 2-Bolton base+CAT"
)

# -----------------------------------------------------------------------------
# TAXA TO EXCLUDE - Add any taxa you want to exclude from analysis
# These will be removed BEFORE calculating relative abundance
# -----------------------------------------------------------------------------
taxa_to_exclude <- c(
  "Homo sapiens"
  # Add more taxa to exclude as needed, e.g.:
  # "Mus musculus",
  # "Other contaminant"
)
```

```{r functions}
# -----------------------------------------------------------------------------
# Function to wrap long text labels
# -----------------------------------------------------------------------------
wrap_labels <- function(x, width = 25) {
  sapply(x, function(label) {
    if (is.na(label)) return(label)
    paste(strwrap(label, width = width), collapse = "\n")
  }, USE.NAMES = FALSE)
}

# -----------------------------------------------------------------------------
# Function to get sample name (with optional custom mapping)
# -----------------------------------------------------------------------------
get_sample_name <- function(filename, use_custom = TRUE, mapping = NULL) {
  # Extract barcode from filename
  barcode <- str_extract(basename(filename), "barcode\\d+")
  
  if (use_custom && !is.null(mapping) && barcode %in% names(mapping)) {
    return(mapping[barcode])
  } else {
    # Default: "Barcode 01" format
    return(str_to_title(str_replace(barcode, "barcode", "Barcode ")))
  }
}

# -----------------------------------------------------------------------------
# Function to read Kraken2 report and extract data at specified taxonomic level
# Now excludes specified taxa and recalculates relative abundance
# -----------------------------------------------------------------------------
read_kraken_report <- function(file_path, tax_rank = "S", 
                                exclude_taxa = NULL,
                                use_custom_names = TRUE,
                                name_mapping = NULL) {
  # Read the file (tab-separated, no header)
  df <- read.delim(
    file_path,
    header = FALSE,
    sep = "\t",
    stringsAsFactors = FALSE,
    col.names = c("percentage", "reads_clade", "reads_direct", "rank", "taxid", "name")
  )
  
  # Clean up names (trim whitespace)
  df$name <- trimws(df$name)
  
  # Get sample name
  sample_name <- get_sample_name(file_path, use_custom_names, name_mapping)
  
  # Get total reads (from root or sum of all)
  total_reads <- df %>%
    filter(rank == "R" | (rank == "U" & name == "unclassified")) %>%
    summarise(total = sum(reads_clade)) %>%
    pull(total)
  
  # If no root found, use sum of unclassified + classified
  if (length(total_reads) == 0 || total_reads == 0) {
    total_reads <- sum(df$reads_clade[df$rank %in% c("U", "R")], na.rm = TRUE)
  }
  
  # Get unclassified reads count
  unclassified_reads <- df %>%
    filter(rank == "U") %>%
    pull(reads_clade)
  
  if (length(unclassified_reads) == 0) unclassified_reads <- 0
  
  # Get reads for excluded taxa (at species level for accurate counting)
  excluded_reads <- 0
  if (!is.null(exclude_taxa) && length(exclude_taxa) > 0) {
    # For species-level exclusion, we need to get the clade reads
    excluded_reads <- df %>%
      filter(name %in% exclude_taxa) %>%
      summarise(total = sum(reads_clade, na.rm = TRUE)) %>%
      pull(total)
    
    if (is.na(excluded_reads)) excluded_reads <- 0
    
    cat("  Sample:", sample_name, "- Excluded", excluded_reads, "reads from:", 
        paste(exclude_taxa, collapse = ", "), "\n")
  }
  
  # Calculate adjusted total (excluding the removed taxa)
  adjusted_total <- total_reads - excluded_reads
  
  # Get data at specified taxonomic level (excluding specified taxa)
  taxa <- df %>%
    filter(rank == tax_rank) %>%
    filter(!(name %in% exclude_taxa)) %>%
    mutate(
      sample = sample_name,
      reads = reads_clade
    ) %>%
    select(sample, name, reads)
  
  # Create unclassified entry
  unclassified <- tibble(
    sample = sample_name,
    name = "Unclassified",
    reads = unclassified_reads
  )
  
  # Combine taxa and unclassified
  result <- bind_rows(taxa, unclassified)
  
  # Recalculate relative abundance based on adjusted total
  # This ensures percentages sum to 100% after exclusion
  result <- result %>%
    group_by(sample) %>%
    mutate(
      # Use the sum of remaining reads for normalization
      total_remaining = sum(reads, na.rm = TRUE),
      relative_abundance = (reads / total_remaining) * 100
    ) %>%
    ungroup() %>%
    select(sample, name, relative_abundance, reads)
  
  return(result)
}

# -----------------------------------------------------------------------------
# Function to prepare plot data with Other category
# Now ensures data sums to 100% and uses custom sample order
# -----------------------------------------------------------------------------
prepare_plot_data <- function(all_data, n_top = 15, custom_sample_order = NULL) {
  # Find top taxa (excluding Unclassified)
  top_taxa <- all_data %>%
    filter(name != "Unclassified") %>%
    group_by(name) %>%
    summarise(total_abundance = sum(relative_abundance), .groups = "drop") %>%
    arrange(desc(total_abundance)) %>%
    head(n_top) %>%
    pull(name)
  
  # Create "Other" category
  plot_data <- all_data %>%
    mutate(
      category = case_when(
        name == "Unclassified" ~ "Unclassified",
        name %in% top_taxa ~ name,
        TRUE ~ "Other"
      )
    ) %>%
    group_by(sample, category) %>%
    summarise(
      relative_abundance = sum(relative_abundance), 
      .groups = "drop"
    ) %>%
    rename(name = category)
  
  # IMPORTANT: Normalize to exactly 100% per sample
  plot_data <- plot_data %>%
    group_by(sample) %>%
    mutate(
      sample_total = sum(relative_abundance),
      relative_abundance = (relative_abundance / sample_total) * 100
    ) %>%
    ungroup() %>%
    select(-sample_total)
  
  # Verify normalization
  check_totals <- plot_data %>%
    group_by(sample) %>%
    summarise(total = sum(relative_abundance), .groups = "drop")
  
  cat("\nVerification - Sample totals (should all be 100%):\n")
  print(check_totals)
  
  # Define category order
  category_order <- c(top_taxa, "Other", "Unclassified")
  
  plot_data <- plot_data %>%
    mutate(name = factor(name, levels = rev(category_order))) %>%
    complete(sample, name, fill = list(relative_abundance = 0))
  
  # Sort samples - use custom order if provided
  if (!is.null(custom_sample_order)) {
    # Use the provided custom order
    sample_order <- custom_sample_order[custom_sample_order %in% unique(plot_data$sample)]
    # Add any samples not in custom order
    remaining <- setdiff(unique(plot_data$sample), sample_order)
    if (length(remaining) > 0) {
      sample_order <- c(sample_order, sort(remaining))
    }
    cat("\nUsing custom sample order:\n")
    for (s in sample_order) cat("  -", s, "\n")
  } else {
    # Default: sort numerically by number in name
    sample_order <- plot_data %>%
      distinct(sample) %>%
      mutate(num = as.numeric(str_extract(sample, "\\d+"))) %>%
      arrange(num, sample) %>%
      pull(sample)
  }
  
  plot_data <- plot_data %>%
    mutate(sample = factor(sample, levels = sample_order))
  
  return(list(data = plot_data, top_taxa = top_taxa, category_order = category_order))
}

# -----------------------------------------------------------------------------
# Function to generate color palette - ENHANCED VERSION
# -----------------------------------------------------------------------------
generate_colors <- function(top_taxa, tax_level = "species") {
  n <- length(top_taxa)
  
  # Define a diverse color palette with distinct, visually separable colors
  # These colors are chosen to be maximally different from each other
  distinct_colors <- c(
    "#1f77b4",  # Blue
    "#ff7f0e",  # Orange
    "#d62728",  # Red
    "#9467bd",  # Purple
    "#8c564b",  # Brown
    "#e377c2",  # Pink
    "#17becf",  # Cyan
    "#bcbd22",  # Yellow-green
    "#7f7f7f",  # Gray
    "#aec7e8",  # Light blue
    "#ffbb78",  # Light orange
    "#ff9896",  # Light red
    "#c5b0d5",  # Light purple
    "#c49c94",  # Tan
    "#f7b6d2",  # Light pink
    "#c7c7c7",  # Light gray
    "#dbdb8d",  # Light yellow-green
    "#9edae5",  # Light cyan
    "#393b79",  # Dark blue
    "#637939"   # Olive
  )
  
  # If we need more colors, repeat the palette
  if (n > length(distinct_colors)) {
    distinct_colors <- rep(distinct_colors, ceiling(n / length(distinct_colors)))
  }
  
  # Create named vector
  colors <- setNames(distinct_colors[1:n], top_taxa)
  
  # Override Campylobacter with greens (KEEP ORIGINAL)
  campylobacter_taxa <- top_taxa[str_detect(top_taxa, regex("Campylobacter", ignore_case = TRUE))]
  green_shades <- c("#33a02c", "#b2df8a", "#006d2c", "#74c476")
  for (i in seq_along(campylobacter_taxa)) {
    colors[campylobacter_taxa[i]] <- green_shades[min(i, length(green_shades))]
  }
  
  # Add Other and Unclassified (KEEP ORIGINAL)
  colors["Other"] <- "#b0b0b0"
  colors["Unclassified"] <- "#404040"
  
  return(colors)
}

# -----------------------------------------------------------------------------
# Function to create stacked bar plot - ENHANCED VERSION
# -----------------------------------------------------------------------------
create_stacked_plot <- function(plot_data, colors, title, label_width = 25) {
  # Create wrapped labels for the legend
  plot_data <- plot_data %>%
    mutate(name_wrapped = wrap_labels(as.character(name), width = label_width))
  
  # Update colors to use wrapped names
  colors_wrapped <- colors
  names(colors_wrapped) <- wrap_labels(names(colors), width = label_width)
  
  # Get the factor order with wrapped names
  name_levels <- levels(plot_data$name)
  wrapped_levels <- wrap_labels(name_levels, width = label_width)
  plot_data$name_wrapped <- factor(plot_data$name_wrapped, levels = wrapped_levels)
  
  ggplot(plot_data, aes(x = sample, y = relative_abundance, fill = name_wrapped)) +
    geom_bar(stat = "identity", position = "stack", width = 0.8, color = "white", linewidth = 0.2) +
    scale_fill_manual(values = colors_wrapped, name = "Taxa") +
    # ENHANCED: More y-axis breaks (every 10%)
    scale_y_continuous(
      expand = c(0, 0), 
      limits = c(0, 100.5),
      breaks = seq(0, 100, by = 10),  # Y-axis breaks every 10%
      minor_breaks = seq(0, 100, by = 5)  # Minor breaks every 5%
    ) +
    labs(
      title = title,
      x = "Sample",
      y = "Relative Abundance (%)"
    ) +
    theme_minimal(base_size = 12) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 11),
      axis.text.y = element_text(size = 10),
      legend.position = "right",
      legend.text = element_text(size = 8, face = "italic"),
      legend.title = element_text(size = 10, face = "bold"),
      legend.key.height = unit(1.2, "lines"),
      plot.title = element_text(face = "bold", size = 14),
      panel.grid.major.x = element_blank(),
      panel.grid.major.y = element_line(color = "gray90", linewidth = 0.5),  # Add horizontal gridlines
      panel.grid.minor.y = element_line(color = "gray95", linewidth = 0.25)  # Add minor gridlines
    ) +
    guides(fill = guide_legend(reverse = TRUE, ncol = 1))
}

# -----------------------------------------------------------------------------
# Function to create heatmap
# -----------------------------------------------------------------------------
create_heatmap <- function(plot_data, category_order, title, label_width = 25) {
  # Create wrapped labels for y-axis
  heatmap_data <- plot_data %>%
    mutate(
      name = factor(name, levels = category_order),
      name_wrapped = wrap_labels(as.character(name), width = label_width)
    )
  
  # Set factor order for wrapped names
  wrapped_order <- wrap_labels(category_order, width = label_width)
  heatmap_data$name_wrapped <- factor(heatmap_data$name_wrapped, levels = wrapped_order)
  
  ggplot(heatmap_data, aes(x = sample, y = name_wrapped, fill = relative_abundance)) +
    geom_tile(color = "white", linewidth = 0.5) +
    geom_text(
      aes(label = ifelse(relative_abundance > 0.5, sprintf("%.1f", relative_abundance), "")),
      size = 2.5,
      color = ifelse(heatmap_data$relative_abundance > 50, "white", "black")
    ) +
    scale_fill_gradient(
      low = "#FFFFCC",
      high = "#800026",
      name = "Relative\nAbundance (%)",
      limits = c(0, 100)
    ) +
    labs(
      title = title,
      x = "Sample",
      y = "Taxa"
    ) +
    theme_minimal(base_size = 12) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 11),
      axis.text.y = element_text(size = 9, face = "italic"),
      legend.position = "right",
      plot.title = element_text(face = "bold", size = 14),
      panel.grid = element_blank()
    )
}
```

# Data Loading

```{r read_files}
# -----------------------------------------------------------------------------
# Read all Kraken2 report files
# -----------------------------------------------------------------------------
files <- list.files(
  path = data_dir,
  pattern = "kraken2\\.report$",
  full.names = TRUE
)

cat("Found", length(files), "Kraken2 report files:\n")
for (f in files) {
  cat("  -", basename(f), "\n")
}

cat("\n")
cat("Taxa being excluded from analysis:\n")
for (t in taxa_to_exclude) {
  cat("  -", t, "\n")
}
cat("\n")

if (use_custom_names) {
  cat("Using custom sample names:\n")
  for (bc in names(sample_name_mapping)) {
    cat("  ", bc, "->", sample_name_mapping[bc], "\n")
  }
} else {
  cat("Using default barcode names\n")
}

cat("\nSample order for plots:\n")
for (s in sample_order_custom) {
  cat("  -", s, "\n")
}
```

# Species Level Analysis

```{r species_analysis, fig.width=14, fig.height=8}
# =============================================================================
# SPECIES LEVEL ANALYSIS
# =============================================================================
cat("\n", paste(rep("=", 70), collapse = ""), "\n")
cat("SPECIES LEVEL ANALYSIS (excluding:", paste(taxa_to_exclude, collapse = ", "), ")\n")
cat(paste(rep("=", 70), collapse = ""), "\n")

# Read species data with exclusions
species_data <- map_dfr(files, ~read_kraken_report(
  .x, 
  tax_rank = "S",
  exclude_taxa = taxa_to_exclude,
  use_custom_names = use_custom_names,
  name_mapping = sample_name_mapping
))

# Prepare plot data
species_prep <- prepare_plot_data(species_data, n_top = n_top_taxa, custom_sample_order = sample_order_custom)
species_plot_data <- species_prep$data
species_top <- species_prep$top_taxa
species_order <- species_prep$category_order

cat("\nTop", n_top_taxa, "Species (after exclusions):\n")
for (i in seq_along(species_top)) {
  total <- sum(species_data$relative_abundance[species_data$name == species_top[i]])
  cat(sprintf("  %2d. %s (%.2f%%)\n", i, species_top[i], total))
}

# Generate colors
species_colors <- generate_colors(species_top, "species")

# Create plots
p_species_stacked <- create_stacked_plot(
  species_plot_data, 
  species_colors,
  paste("Relative Abundance of Top", n_top_taxa, "Bacterial Species"),
  label_width = label_wrap_width
)

p_species_heatmap <- create_heatmap(
  species_plot_data,
  species_order,
  paste("Heatmap of Top", n_top_taxa, "Bacterial Species"),
  label_width = label_wrap_width
)

# Display plots
print(p_species_stacked)
print(p_species_heatmap)

# Save plots
ggsave("species_abundance_stacked.png", plot = p_species_stacked, width = 14, height = 8, dpi = 300)
ggsave("species_abundance_heatmap.png", plot = p_species_heatmap, width = 12, height = 10, dpi = 300)
cat("\nSaved: species_abundance_stacked.png\n")
cat("Saved: species_abundance_heatmap.png\n")
```

# Genus Level Analysis

```{r genus_analysis, fig.width=14, fig.height=8}
# =============================================================================
# GENUS LEVEL ANALYSIS
# =============================================================================
cat("\n", paste(rep("=", 70), collapse = ""), "\n")
cat("GENUS LEVEL ANALYSIS\n")
cat(paste(rep("=", 70), collapse = ""), "\n")

# For genus level, we exclude "Homo" instead of "Homo sapiens"
genus_exclude <- c("Homo")

# Read genus data
genus_data <- map_dfr(files, ~read_kraken_report(
  .x, 
  tax_rank = "G",
  exclude_taxa = genus_exclude,
  use_custom_names = use_custom_names,
  name_mapping = sample_name_mapping
))

# Prepare plot data
genus_prep <- prepare_plot_data(genus_data, n_top = n_top_taxa, custom_sample_order = sample_order_custom)
genus_plot_data <- genus_prep$data
genus_top <- genus_prep$top_taxa
genus_order <- genus_prep$category_order

cat("\nTop", n_top_taxa, "Genera:\n")
for (i in seq_along(genus_top)) {
  total <- sum(genus_data$relative_abundance[genus_data$name == genus_top[i]])
  cat(sprintf("  %2d. %s (%.2f%%)\n", i, genus_top[i], total))
}

# Generate colors
genus_colors <- generate_colors(genus_top, "genus")

# Create plots
p_genus_stacked <- create_stacked_plot(
  genus_plot_data, 
  genus_colors,
  paste("Relative Abundance of Top", n_top_taxa, "Bacterial Genera"),
  label_width = label_wrap_width
)

p_genus_heatmap <- create_heatmap(
  genus_plot_data,
  genus_order,
  paste("Heatmap of Top", n_top_taxa, "Bacterial Genera"),
  label_width = label_wrap_width
)

# Display plots
print(p_genus_stacked)
print(p_genus_heatmap)

# Save plots
ggsave("genus_abundance_stacked.png", plot = p_genus_stacked, width = 14, height = 8, dpi = 300)
ggsave("genus_abundance_heatmap.png", plot = p_genus_heatmap, width = 12, height = 10, dpi = 300)
cat("\nSaved: genus_abundance_stacked.png\n")
cat("Saved: genus_abundance_heatmap.png\n")
```

# Family Level Analysis

```{r family_analysis, fig.width=14, fig.height=8}
# =============================================================================
# FAMILY LEVEL ANALYSIS
# =============================================================================
cat("\n", paste(rep("=", 70), collapse = ""), "\n")
cat("FAMILY LEVEL ANALYSIS\n")
cat(paste(rep("=", 70), collapse = ""), "\n")

# For family level, exclude "Hominidae" instead of "Homo sapiens"
family_exclude <- c("Hominidae")

# Read family data
family_data <- map_dfr(files, ~read_kraken_report(
  .x, 
  tax_rank = "F",
  exclude_taxa = family_exclude,
  use_custom_names = use_custom_names,
  name_mapping = sample_name_mapping
))

# Prepare plot data
family_prep <- prepare_plot_data(family_data, n_top = n_top_taxa, custom_sample_order = sample_order_custom)
family_plot_data <- family_prep$data
family_top <- family_prep$top_taxa
family_order <- family_prep$category_order

cat("\nTop", n_top_taxa, "Families:\n")
for (i in seq_along(family_top)) {
  total <- sum(family_data$relative_abundance[family_data$name == family_top[i]])
  cat(sprintf("  %2d. %s (%.2f%%)\n", i, family_top[i], total))
}

# Generate colors
family_colors <- generate_colors(family_top, "family")

# Create plots
p_family_stacked <- create_stacked_plot(
  family_plot_data, 
  family_colors,
  paste("Relative Abundance of Top", n_top_taxa, "Bacterial Families"),
  label_width = label_wrap_width
)

p_family_heatmap <- create_heatmap(
  family_plot_data,
  family_order,
  paste("Heatmap of Top", n_top_taxa, "Bacterial Families"),
  label_width = label_wrap_width
)

# Display plots
print(p_family_stacked)
print(p_family_heatmap)

# Save plots
ggsave("family_abundance_stacked.png", plot = p_family_stacked, width = 14, height = 8, dpi = 300)
ggsave("family_abundance_heatmap.png", plot = p_family_heatmap, width = 12, height = 10, dpi = 300)
cat("\nSaved: family_abundance_stacked.png\n")
cat("Saved: family_abundance_heatmap.png\n")
```

# Summary Tables

```{r summary_tables}
# =============================================================================
# SAVE SUMMARY TABLES
# =============================================================================

# Species summary
species_summary <- species_plot_data %>%
  pivot_wider(names_from = name, values_from = relative_abundance) %>%
  arrange(sample)
write_csv(species_summary, "species_abundance_summary.csv")

# Genus summary
genus_summary <- genus_plot_data %>%
  pivot_wider(names_from = name, values_from = relative_abundance) %>%
  arrange(sample)
write_csv(genus_summary, "genus_abundance_summary.csv")

# Family summary
family_summary <- family_plot_data %>%
  pivot_wider(names_from = name, values_from = relative_abundance) %>%
  arrange(sample)
write_csv(family_summary, "family_abundance_summary.csv")

cat("\nSummary tables saved:\n")
cat("  - species_abundance_summary.csv\n")
cat("  - genus_abundance_summary.csv\n")
cat("  - family_abundance_summary.csv\n")
```

## Species Summary Table

```{r print_species_summary}
cat("\n", paste(rep("=", 70), collapse = ""), "\n")
cat("SPECIES SUMMARY (Homo sapiens excluded)\n")
cat(paste(rep("=", 70), collapse = ""), "\n")
print(species_summary, n = Inf, width = Inf)
```

## Genus Summary Table

```{r print_genus_summary}
cat("\n", paste(rep("=", 70), collapse = ""), "\n")
cat("GENUS SUMMARY (Homo excluded)\n")
cat(paste(rep("=", 70), collapse = ""), "\n")
print(genus_summary, n = Inf, width = Inf)
```

## Family Summary Table

```{r print_family_summary}
cat("\n", paste(rep("=", 70), collapse = ""), "\n")
cat("FAMILY SUMMARY (Hominidae excluded)\n")
cat(paste(rep("=", 70), collapse = ""), "\n")
print(family_summary, n = Inf, width = Inf)

cat("\n✓ All outputs generated successfully!\n")
cat("\nNote: Relative abundances have been recalculated after excluding human reads.\n")
cat("All bars should now sum to exactly 100%.\n")
```

---

## Key Enhancements in This Version

1. **Enhanced Y-axis**: Y-axis now shows tick marks at every 10% (0, 10, 20, ... 100) with minor gridlines every 5%

2. **Improved Color Palette**: 
   - 20 distinct, maximally different colors for better visual separation
   - Campylobacter species retain their original green colors
   - "Other" and "Unclassified" categories retain their original gray tones
   
3. **Better Readability**: 
   - Horizontal gridlines added to make reading exact percentages easier
   - White borders between stacked bars for clearer separation

4. **HTML Output**: This R Markdown file generates an HTML report with:
   - Table of contents for easy navigation
   - Code folding option
   - Professional theme (Cosmo)
   - All plots and tables embedded

## How to Use

1. Update the `data_dir` variable in the setup chunk to point to your data directory
2. Customize sample names in the `sample_name_mapping` if needed
3. Knit the document to HTML: `rmarkdown::render("bacterial_abundance_analysis_enhanced.Rmd")`
4. Or use the "Knit" button in RStudio
